Supporting Information

### 2018-03-06 SupportingInformation.Rmd
in the midst of adding asymmetry of tail dependence results to the C/N analysis

### Non-paramtric analysis of tail dependence
<!--Do the intensive part of the analysis-->
```{r stat_soilCN_graphic_approach, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
set.seed(seed)

@@

#utility function for calculating fractions of surrogates 
#with a stat smaller than that of data, i.e., fraction of 
#surrogates for which the stat value on the data is bigger
fracwork<-function(dvals,surrvals)
{
  frac<-NA*numeric(length(corstats_d))
  names(frac)<-names(dvals)
  for (counter in 1:length(frac))
  {
    frac[counter]<-sum(surrvals[counter,]<dvals[counter])
  }
  return(frac)
}

#now work with the correlation stats
corstats_d<-calcstats(v_CN,Corbds,"Cor")

corstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Corbds,nm="Cor")
corstats_frac_K<-fracwork(corstats_d,corstats_K)
corstats_K<-apply(FUN=quantile,X=corstats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))

corstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Corbds,nm="Cor")
corstats_frac_S<-fracwork(corstats_d,corstats_S)
corstats_S<-apply(FUN=quantile,X=corstats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))

#now work with the D2 stats
D2stats_d<-calcstats(v_CN,D2bds,"Dtwo")

D2stats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=D2bds,nm="Dtwo")
D2stats_frac_K<-fracwork(D2stats_d,D2stats_K)
D2stats_K<-apply(FUN=quantile,X=D2stats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))

D2stats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=D2bds,nm="Dtwo")
D2stats_frac_S<-fracwork(D2stats_d,D2stats_S)
D2stats_S<-apply(FUN=quantile,X=D2stats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))

@@

Pstats_d<-calcstats(v_CN,Pbds_wrap,"P")

Pstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Pbds_wrap,nm="P")
Pstats_frac_K<-fracwork(Pstats_d,Pstats_K)
Pstats_K<-apply(FUN=quantile,X=Pstats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))

Pstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Pbds_wrap,nm="P")
Pstats_frac_S<-fracwork(Pstats_d,Pstats_S)
Pstats_S<-apply(FUN=quantile,X=Pstats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
```

<!--Now do the plotting-->
```{r plot_soilCN_graphic_approach, echo=F}
x<-seq(from=0.05,to=0.95,by=0.1)
xlimits<-c(0,1)

#plotting layout, units inches
xht<-0.5
ywd<-0.5
titleht<-.25
panht<-1.5
panwd<-3
gap<-.05
totwd<-ywd+2*panwd+2*gap
totht<-xht+3*panht+3*gap+titleht
pdf(file="CNdata_NonparamTailDep.pdf",width=totwd,height=totht)

#plot kendall results on the left panels, spearman on right,
#cor in top panels, then D2, then P on bottom panels

@@

#kendall, D2
par(fig=c(ywd/totwd,
          (ywd+panwd)/totwd,
          (xht+panht+gap)/totht,
          (xht+2*panht+gap)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
ylimits_D2<-range(D2stats_d,D2stats_K,D2stats_K)
ylimits_D2[2]<-ylimits_D2[2]+.3*diff(ylimits_D2)
plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2,
     xaxt='n')
mtext(side=2,line=1,text=expression(D[2]))
axis(side=1,labels=F)
lines(x,D2stats_K[1,],type='l',lty='dotted')
lines(x,D2stats_K[2,],type='l',lty='dashed')
lines(x,D2stats_K[3,],type='l',lty='dashed')
lines(x,D2stats_K[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_D2[1],labels='C',cex=1.5,adj=c(.5,0))
fracplot(D2stats_frac_K,ylimits_D2)
vlineplot(x,ylimits_D2)

#spearman, D2
par(fig=c((ywd+panwd+gap)/totwd,
          (ywd+2*panwd+gap)/totwd,
          (xht+panht+gap)/totht,
          (xht+2*panht+gap)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2,
     xaxt='n',yaxt='n')
axis(side=1,labels=F)
axis(side=2,labels=F)
lines(x,D2stats_S[1,],type='l',lty='dotted')
lines(x,D2stats_S[2,],type='l',lty='dashed')
lines(x,D2stats_S[3,],type='l',lty='dashed')
lines(x,D2stats_S[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_D2[1],labels='D',cex=1.5,adj=c(.5,0))
fracplot(D2stats_frac_S,ylimits_D2)
vlineplot(x,ylimits_D2)

#kendall, P
par(fig=c(ywd/totwd,
          (ywd+panwd)/totwd,
          (xht)/totht,
          (xht+panht)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
ylimits_P<-range(Pstats_d,Pstats_K,Pstats_K)
ylimits_P[2]<-ylimits_P[2]+.3*diff(ylimits_P)
plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P)
mtext(side=2,line=1,text="P")
mtext(side=1,line=1,text="Diagonal slice")
lines(x,Pstats_K[1,],type='l',lty='dotted')
lines(x,Pstats_K[2,],type='l',lty='dashed')
lines(x,Pstats_K[3,],type='l',lty='dashed')
lines(x,Pstats_K[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_P[1],labels='E',cex=1.5,adj=c(.5,0))
fracplot(Pstats_frac_K,ylimits_P)
vlineplot(x,ylimits_P)

#spearman, P
par(fig=c((ywd+panwd+gap)/totwd,
          (ywd+2*panwd+gap)/totwd,
          (xht)/totht,
          (xht+panht)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,
     yaxt='n')
mtext(side=1,line=1,text="Diagonal slice")
axis(side=2,labels=F)
lines(x,Pstats_S[1,],type='l',lty='dotted')
lines(x,Pstats_S[2,],type='l',lty='dashed')
lines(x,Pstats_S[3,],type='l',lty='dashed')
lines(x,Pstats_S[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_P[1],labels='F',cex=1.5,adj=c(.5,0))
fracplot(Pstats_frac_S,ylimits_P)
vlineplot(x,ylimits_P)

dev.off()
#DAN: have to make the plot actually appear in the Rmd output, with a caption and everything!
```

>>>>

### Non-parametric analysis of tail dependence
<!--Do the intensive part of the analysis, calculating stats on the data and surrogates-->
```{r stat_soilCN_graphic_approach, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
set.seed(seed)

@@

#utility function for calculating fractions of surrogates 
#with a stat smaller than that of data, i.e., fraction of 
#surrogates for which the stat value on the data is bigger
fracwork<-function(dvals,surrvals)
{
  frac<-NA*numeric(length(dvals))
  names(frac)<-names(dvals)
  for (counter in 1:length(frac))
  {
    frac[counter]<-sum(surrvals[counter,]<dvals[counter])
  }
  return(frac)
}

#now work with the correlation stats
corstats_d<-calcstats(v_CN,Corbds,"Cor")
corstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Corbds,nm="Cor")
corstats_frac_K<-fracwork(corstats_d,corstats_K)

corlmcoru_d<-unname(corstats_d[1]-corstats_d[length(corstats_d)])
corlmcoru_K<-corstats_K[,1]-corstats_K[,numsurrog]
corlmcoru_frac_K<-sum(corlmcoru_K<corlmcoru_d)
corstats_K<-apply(FUN=quantile,X=corstats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))

corstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Corbds,nm="Cor")
corstats_frac_S<-fracwork(corstats_d,corstats_S)

corlmcoru_S<-corstats_S[,1]-corstats_S[,numsurrog]
corlmcoru_frac_S<-sum(corlmcoru_S<corlmcoru_d)
corstats_S<-apply(FUN=quantile,X=corstats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))

@@

Pstats_d<-calcstats(v_CN,Pbds_wrap,"P")
Pstats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=Pbds_wrap,nm="P")
Pstats_frac_K<-fracwork(Pstats_d,Pstats_K)

PlmPu_d<-unname(Pstats_d[1]-Pstats_d[length(Pstats_d)])
PlmPu_K<-Pstats_K[,1]-Pstats_K[,numsurrog]
PlmPu_frac_K<-sum(PlmPu_K<PlmPu_d)
Pstats_K<-apply(FUN=quantile,X=Pstats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))

Pstats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=Pbds_wrap,nm="P")
Pstats_frac_S<-fracwork(Pstats_d,Pstats_S)

PlmPu_S<-Pstats_S[,1]-Pstats_S[,numsurrog]
PlmPu_frac_S<-sum(PlmPu_S<PlmPu_d)
Pstats_S<-apply(FUN=quantile,X=Pstats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))

#now work with the D2 stats
D2stats_d<-calcstats(v_CN,D2bds,"Dtwo")
D2stats_K<-apply(FUN=calcstats,X=surv_K,MARGIN=3,f=D2bds,nm="Dtwo")
D2stats_frac_K<-fracwork(D2stats_d,D2stats_K)

D2lmD2u_d<-D2stats_d[1]-D2stats_d[length(D2stats_d)]
D2lmD2u_K<-D2stats_K[,1]-D2stats_K[,numsurrog]
D2lmD2u_frac_K<-sum(D2lmD2u_K<D2lmD2u_d)
D2stats_K<-apply(FUN=quantile,X=D2stats_K,MARGIN=1,prob=c(.005,0.25,.975,.995))

D2stats_S<-apply(FUN=calcstats,X=surv_S,MARGIN=3,f=D2bds,nm="Dtwo")
D2stats_frac_S<-fracwork(D2stats_d,D2stats_S)

D2lmD2u_S<-D2stats_S[,1]-D2stats_S[,numsurrog]
D2lmD2u_frac_S<-sum(D2lmD2u_S<D2lmD2u_d)
D2stats_S<-apply(FUN=quantile,X=D2stats_S,MARGIN=1,prob=c(.005,0.25,.975,.995))
```

<!--Now save the plot to the results folder-->
```{r plot_soilCN_graphic_approach, echo=F}
x<-seq(from=0.05,to=0.95,by=0.1)
xlimits<-c(0,1)

#plotting layout, units inches
xht<-0.5
ywd<-0.5
titleht<-.25
panht<-1.5
panwd<-3
gap<-.05
totwd<-ywd+2*panwd+2*gap
totht<-xht+3*panht+3*gap+titleht
pdf(file="./Results/stat_results/stat_soilCN//CNdata_NonparamTailDep.pdf",width=totwd,height=totht)

#plot kendall results on the left panels, spearman on right,
#cor in top panels, then P, then D2 on bottom panels

@@

#kendall, P
par(fig=c(ywd/totwd,
          (ywd+panwd)/totwd,
          (xht+panht+gap)/totht,
          (xht+2*panht+gap)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
ylimits_D2<-range(D2stats_d,D2stats_K,D2stats_K)
ylimits_D2[2]<-ylimits_D2[2]+.3*diff(ylimits_D2)
ylimits_P<-range(Pstats_d,Pstats_K,Pstats_K)
ylimits_P[2]<-ylimits_P[2]+.3*diff(ylimits_P)
plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,xaxt='n')
axis(side=1,labels=F)
mtext(side=2,line=1,text="P")
lines(x,Pstats_K[1,],type='l',lty='dotted')
lines(x,Pstats_K[2,],type='l',lty='dashed')
lines(x,Pstats_K[3,],type='l',lty='dashed')
lines(x,Pstats_K[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_P[1],labels='C',cex=1.5,adj=c(.5,0))
fracplot(Pstats_frac_K,ylimits_P)
vlineplot(x,ylimits_P)

#spearman, p
par(fig=c((ywd+panwd+gap)/totwd,
          (ywd+2*panwd+gap)/totwd,
          (xht+panht+gap)/totht,
          (xht+2*panht+gap)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,
     xaxt='n',yaxt='n')
axis(side=1,labels=F)
axis(side=2,labels=F)
lines(x,Pstats_S[1,],type='l',lty='dotted')
lines(x,Pstats_S[2,],type='l',lty='dashed')
lines(x,Pstats_S[3,],type='l',lty='dashed')
lines(x,Pstats_S[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_P[1],labels='D',cex=1.5,adj=c(.5,0))
fracplot(Pstats_frac_S,ylimits_P)
vlineplot(x,ylimits_P)

#kendall, D2
par(fig=c(ywd/totwd,
          (ywd+panwd)/totwd,
          (xht)/totht,
          (xht+panht)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2)
mtext(side=1,line=1,text="Diagonal slice")
mtext(side=2,line=1,text=expression(D[2]))
axis(side=1,labels=F)
lines(x,D2stats_K[1,],type='l',lty='dotted')
lines(x,D2stats_K[2,],type='l',lty='dashed')
lines(x,D2stats_K[3,],type='l',lty='dashed')
lines(x,D2stats_K[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_D2[1],labels='E',cex=1.5,adj=c(.5,0))
fracplot(D2stats_frac_K,ylimits_D2)
vlineplot(x,ylimits_D2)

#spearman, P
par(fig=c((ywd+panwd+gap)/totwd,
          (ywd+2*panwd+gap)/totwd,
          (xht)/totht,
          (xht+panht)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2,
     yaxt='n')
mtext(side=1,line=1,text="Diagonal slice")
axis(side=1,labels=F)
axis(side=2,labels=F)
lines(x,D2stats_S[1,],type='l',lty='dotted')
lines(x,D2stats_S[2,],type='l',lty='dashed')
lines(x,D2stats_S[3,],type='l',lty='dashed')
lines(x,D2stats_S[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_D2[1],labels='F',cex=1.5,adj=c(.5,0))
fracplot(D2stats_frac_S,ylimits_D2)
vlineplot(x,ylimits_D2)

dev.off()
```

<!--Now include the plot in the Rmd output-->
```{r show_plot_soildCN,echo=F}
fig_nums("soilCN_nonparam",display=F)
```
![Nonparametric tests for tail dependence and other deviations from a normal copula for soil C and N data. As described in Methods, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles).  When red lines were outside dashed lines, text at the top of plots indicates how many surrogate values the data value was less than or greater than.](./Results/stat_results/stat_soilCN//CNdata_NonparamTailDep.pdf)

<!--Now save a second plot to the results folder, this one about asymmetry of tail dependence-->
```{r plot_soilCN_graphic_approach_2, echo=F}
#plotting layout, units inches
xht<-0.5
ywd<-0.5
titleht<-.25
panht<-1.25
panwd<-3
gap<-.05
totwd<-2*ywd+2*panwd+2*gap
totht<-3*xht+3*panht+3*gap+titleht
pdf(file="./Results/stat_results/stat_soilCN//CNdata_AsymmTailDep.pdf",width=totwd,height=totht)

#plot kendall results on the left panels, spearman on right,
#cor in top panels, then P, then D2 on bottom panels

#kendall, cor
par(fig=c(ywd/totwd,
          (ywd+panwd)/totwd,
          (3*xht+2*panht+2*gap)/totht,
          (3*xht+3*panht+2*gap)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
hist(corlmcoru_K,50)
points(corlmcoru_d,0,col='red')
mtext(side=3,line=0,text="Kendall-preserving surrogates")
mtext(side=1,line=1,text="Partial correlation")
#text(xlimits[1],ylimits_cor[1],labels='A',cex=1.5,adj=c(.5,0))
fracplot<-function(fracs,ylims)
{
  for (counter in 1:length(fracs))
  {
    ptxt<-''
    if (fracs[counter]>.975*numsurrog)
    {
      ptxt<-paste(">",fracs[counter],sep='')
    } 
    if (fracs[counter]<.025*numsurrog)
    {
      ptxt<-paste("<",numsurrog-fracs[counter],sep='')
    }
    yht<-ylims[2]-.1*diff(ylims)
    text(x[counter],yht,ptxt,adj=c(0.5,.5),cex=0.5,srt=90)
  }
}
fracplot(corstats_frac_K,ylimits_cor)
vlineplot<-function(x,ylims)
{
  for (counter in 1:length(x))
  {
    lines(rep(x[counter]-(x[2]-x[1])/2,2),ylims,type='l',lty='dotted')
  }
  lines(rep(x[length(x)]+(x[2]-x[1])/2,2),ylims,type='l',lty='dotted')
}
vlineplot(x,ylimits_cor)

#kendall, P
par(fig=c(ywd/totwd,
          (ywd+panwd)/totwd,
          (xht+panht+gap)/totht,
          (xht+2*panht+gap)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
ylimits_D2<-range(D2stats_d,D2stats_K,D2stats_K)
ylimits_D2[2]<-ylimits_D2[2]+.3*diff(ylimits_D2)
ylimits_P<-range(Pstats_d,Pstats_K,Pstats_K)
ylimits_P[2]<-ylimits_P[2]+.3*diff(ylimits_P)
plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,xaxt='n')
axis(side=1,labels=F)
mtext(side=2,line=1,text="P")
lines(x,Pstats_K[1,],type='l',lty='dotted')
lines(x,Pstats_K[2,],type='l',lty='dashed')
lines(x,Pstats_K[3,],type='l',lty='dashed')
lines(x,Pstats_K[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_P[1],labels='C',cex=1.5,adj=c(.5,0))
fracplot(Pstats_frac_K,ylimits_P)
vlineplot(x,ylimits_P)

#spearman, p
par(fig=c((ywd+panwd+gap)/totwd,
          (ywd+2*panwd+gap)/totwd,
          (xht+panht+gap)/totht,
          (xht+2*panht+gap)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
plot(x,Pstats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_P,
     xaxt='n',yaxt='n')
axis(side=1,labels=F)
axis(side=2,labels=F)
lines(x,Pstats_S[1,],type='l',lty='dotted')
lines(x,Pstats_S[2,],type='l',lty='dashed')
lines(x,Pstats_S[3,],type='l',lty='dashed')
lines(x,Pstats_S[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_P[1],labels='D',cex=1.5,adj=c(.5,0))
fracplot(Pstats_frac_S,ylimits_P)
vlineplot(x,ylimits_P)

#kendall, D2
par(fig=c(ywd/totwd,
          (ywd+panwd)/totwd,
          (xht)/totht,
          (xht+panht)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2)
mtext(side=1,line=1,text="Diagonal slice")
mtext(side=2,line=1,text=expression(D[2]))
axis(side=1,labels=F)
lines(x,D2stats_K[1,],type='l',lty='dotted')
lines(x,D2stats_K[2,],type='l',lty='dashed')
lines(x,D2stats_K[3,],type='l',lty='dashed')
lines(x,D2stats_K[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_D2[1],labels='E',cex=1.5,adj=c(.5,0))
fracplot(D2stats_frac_K,ylimits_D2)
vlineplot(x,ylimits_D2)

#spearman, P
par(fig=c((ywd+panwd+gap)/totwd,
          (ywd+2*panwd+gap)/totwd,
          (xht)/totht,
          (xht+panht)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2,
     yaxt='n')
mtext(side=1,line=1,text="Diagonal slice")
axis(side=1,labels=F)
axis(side=2,labels=F)
lines(x,D2stats_S[1,],type='l',lty='dotted')
lines(x,D2stats_S[2,],type='l',lty='dashed')
lines(x,D2stats_S[3,],type='l',lty='dashed')
lines(x,D2stats_S[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_D2[1],labels='F',cex=1.5,adj=c(.5,0))
fracplot(D2stats_frac_S,ylimits_D2)
vlineplot(x,ylimits_D2)

dev.off()
```
